%% ──────────────────────────────────────────────
%%  کتاب‌نامه
%%  فایل: backmatter/bibliography.tex
%% ──────────────────────────────────────────────

\chapter*{کتاب‌نامه}
\addcontentsline{toc}{chapter}{کتاب‌نامه}
\label{ch:bibliography}

\begin{tcolorbox}[mybox, title={درباره‌ی کتاب‌نامه}]
منابع زیر شامل تمامی آثار ارجاع‌داده‌شده در ۱۴ فصل و ۴ پیوست این کتاب است. منابع به سه بخش تفکیک شده‌اند: فارسی، انگلیسی، و اسناد رسمی/بین‌المللی. منابع انگلیسی به ترتیب الفبایی نام خانوادگی نویسنده و منابع فارسی به ترتیب الفبایی فارسی مرتب شده‌اند.
\end{tcolorbox}

%% ================================================================
\section*{الف) منابع فارسی}
\addcontentsline{toc}{section}{منابع فارسی}
%% ================================================================

\begin{enumerate}[label={[\arabic*]}, leftmargin=2cm, itemsep=6pt]

\item آبراهامیان، یرواند. \textit{ایران بین دو انقلاب}. ترجمه‌ی احمد گل‌محمدی و محمدابراهیم فتاحی. تهران: نشر نی، ۱۳۷۷.

\item آبراهامیان، یرواند. \textit{کودتا: ۲۸ مرداد، سیا و ریشه‌های روابط ایران و آمریکا در عصر مدرن}. ترجمه‌ی محمدابراهیم فتاحی. تهران: نشر نی، ۱۳۹۲.

\item بشیریه، حسین. \textit{دیباچه‌ای بر جامعه‌شناسی سیاسی ایران: دوره‌ی جمهوری اسلامی}. تهران: نشر نگاه معاصر، ۱۳۸۱.

\item زرین‌کوب، عبدالحسین. \textit{دو قرن سکوت}. تهران: انتشارات سخن، ۱۳۳۰.

\item زرین‌کوب، عبدالحسین. \textit{تاریخ مردم ایران}. ۲ جلد. تهران: انتشارات امیرکبیر، ۱۳۶۷.

\item طبری، احسان. \textit{جامعه‌ی ایران در دوره‌ی رضاشاه}. تهران: انتشارات حزب توده‌ی ایران، ۱۳۶۰.

\item فوران، جان. \textit{مقاومت شکننده: تاریخ تحولات اجتماعی ایران از سال ۱۵۰۰ میلادی تا انقلاب}. ترجمه‌ی احمد تدین. تهران: مؤسسه‌ی خدمات فرهنگی رسا، ۱۳۷۷.

\item کاتوزیان، محمدعلی همایون. \textit{اقتصاد سیاسی ایران: از مشروطیت تا پایان سلسله‌ی پهلوی}. ترجمه‌ی محمدرضا نفیسی و کامبیز عزیزی. تهران: نشر مرکز، ۱۳۷۲.

\item کدی، نیکی آر. \textit{ریشه‌های انقلاب ایران}. ترجمه‌ی عبدالرحیم گواهی. تهران: دفتر نشر فرهنگ اسلامی، ۱۳۶۹.

\item گازیوروسکی، مارک و مالکوم برن (ویراستاران). \textit{محمد مصدق و کودتای ۲۸ مرداد}. ترجمه‌ی علی مرشدی‌زاد. تهران: نشر قصیده‌سرا، ۱۳۸۴.

\item مجتهدزاده، پیروز. \textit{جغرافیای سیاسی و سیاست جغرافیایی}. تهران: انتشارات سمت، ۱۳۸۱.

\item میلانی، عباس. \textit{نگاهی به شاه}. تورنتو: نشر پرشین سیرکل، ۲۰۱۲.

\end{enumerate}


%% ================================================================
\section*{ب) منابع انگلیسی}
\addcontentsline{toc}{section}{منابع انگلیسی}
%% ================================================================

\begin{enumerate}[label={[\arabic*]}, leftmargin=2cm, itemsep=6pt, resume]

%% ── A ──
\item \lr{Abramson, Lyn Y., Martin E.P. Seligman, and John D. Teasdale. "Learned Helplessness in Humans: Critique and Reformulation," \textit{Journal of Abnormal Psychology}, Vol. 87, No. 1, 1978, pp. 49–74.}

\item \lr{Acemoglu, Daron and James A. Robinson. \textit{Why Nations Fail: The Origins of Power, Prosperity, and Poverty}. New York: Crown Business, 2012.}

\item \lr{Acemoglu, Daron, Simon Johnson, and James A. Robinson. "An African Success Story: Botswana," in Dani Rodrik (ed.), \textit{In Search of Prosperity}. Princeton: Princeton University Press, 2003.}

\item \lr{Ajami, Fouad. \textit{The Arab Predicament: Arab Political Thought and Practice since 1967}. Updated ed., Cambridge: Cambridge University Press, 1992.}

\item \lr{Allawi, Ali A. \textit{The Occupation of Iraq: Winning the War, Losing the Peace}. New Haven: Yale University Press, 2007.}

\item \lr{Amsden, Alice H. \textit{Asia's Next Giant: South Korea and Late Industrialization}. New York: Oxford University Press, 1989.}

\item \lr{Armstrong, Karen. "Wahhabism to ISIS: How Saudi Arabia Exported the Main Source of Global Terrorism," \textit{New Statesman}, 27 November 2014.}

\item \lr{Auty, Richard M. \textit{Sustaining Development in Mineral Economies: The Resource Curse Thesis}. London: Routledge, 1993.}

%% ── B ──
\item \lr{Bacevich, Andrew J. \textit{America's War for the Greater Middle East: A Military History}. New York: Random House, 2016.}

\item \lr{Badie, Dina. "Groupthink, Iraq, and the War on Terror: Explaining US Policy Shift toward Iraq," \textit{Foreign Policy Analysis}, Vol. 6, No. 4, 2010, pp. 277–296.}

\item \lr{Bail, Christopher A., et al. "Exposure to Opposing Views on Social Media Can Increase Political Polarization," \textit{Proceedings of the National Academy of Sciences}, Vol. 115, No. 37, 2018, pp. 9216–9221.}

\item \lr{Bandura, Albert. "Exercise of Human Agency Through Collective Efficacy," \textit{Current Directions in Psychological Science}, Vol. 9, No. 3, 2000, pp. 75–78.}

\item \lr{Bar-Tal, Daniel, et al. "A Sense of Self-Perceived Collective Victimhood in Intractable Conflicts," \textit{International Review of the Red Cross}, Vol. 91, No. 874, 2009, pp. 229–258.}

\item \lr{Bastardi, Anthony, Eric Luis Uhlmann, and Lee Ross. "Wishful Thinking: Belief, Desire, and the Motivated Evaluation of Scientific Evidence," \textit{Psychological Science}, Vol. 22, No. 6, 2011, pp. 731–732.}

\item \lr{Beblawi, Hazem and Giacomo Luciani (eds.). \textit{The Rentier State}. London: Croom Helm, 1987.}

\item \lr{Behrman, Greg. \textit{The Most Noble Adventure: The Marshall Plan and How America Helped Rebuild Europe}. New York: Free Press, 2007.}

\item \lr{Bellamy, Alex J. \textit{The Responsibility to Protect: A Defense}. Oxford: Oxford University Press, 2015.}

\item \lr{Benner, Thorsten. "Brazil as a Norm Entrepreneur: The 'Responsibility While Protecting' Initiative," \textit{Global Public Policy Institute Working Paper}, 2013.}

\item \lr{Bishara, Marwan. \textit{The Invisible Arab: The Promise and Peril of the Arab Revolutions}. New York: Nation Books, 2012.}

\item \lr{Blass, Thomas. \textit{The Man Who Shocked the World: The Life and Legacy of Stanley Milgram}. New York: Basic Books, 2004.}

\item \lr{Blix, Hans. \textit{Disarming Iraq}. New York: Pantheon Books, 2004.}

\item \lr{Blum, William. \textit{Rogue State: A Guide to the World's Only Superpower}. 3rd ed., Monroe, ME: Common Courage Press, 2005.}

\item \lr{Bowden, Mark. \textit{Black Hawk Down: A Story of Modern War}. New York: Atlantic Monthly Press, 1999.}

\item \lr{Bowden, Mark. \textit{Guests of the Ayatollah: The Iran Hostage Crisis}. New York: Grove Press, 2006.}

\item \lr{Bremer, L. Paul III. \textit{My Year in Iraq: The Struggle to Build a Future of Hope}. New York: Simon \& Schuster, 2006.}

\item \lr{Brooks, Rosa. \textit{How Everything Became War and the Military Became Everything}. New York: Simon \& Schuster, 2016.}

\item \lr{Burnham, Gilbert, et al. "Mortality after the 2003 invasion of Iraq: a cross-sectional cluster sample survey," \textit{The Lancet}, Vol. 368, No. 9545, 2006, pp. 1421–1428.}

%% ── C ──
\item \lr{Carroll, William K. (ed.). \textit{Organizing the 1\%: How Corporate Power Works}. Halifax: Fernwood Publishing, 2017.}

\item \lr{Césaire, Aimé. \textit{Discourse on Colonialism}. Trans. Joan Pinkham. New York: Monthly Review Press, 1972 (orig. 1950).}

\item \lr{Chancel, Lucas, Thomas Piketty, et al. \textit{World Inequality Report 2022}. World Inequality Lab, 2021.}

\item \lr{Charap, Samuel and Timothy J. Colton. \textit{Everyone Loses: The Ukraine Crisis and the Ruinous Contest for Post-Soviet Eurasia}. London: Routledge, 2017.}

\item \lr{Chatterjee, Pratap. \textit{Halliburton's Army: How a Well-Connected Texas Oil Company Revolutionized the Way America Makes War}. New York: Nation Books, 2009.}

\item \lr{Chayes, Sarah. \textit{Thieves of State: Why Corruption Threatens Global Security}. New York: W.W. Norton, 2015.}

\item \lr{Chenoweth, Erica and Maria J. Stephan. \textit{Why Civil Resistance Works: The Strategic Logic of Nonviolent Conflict}. New York: Columbia University Press, 2011.}

\item \lr{Chollet, Derek. \textit{The Long Game: How Obama Defied Washington and Redefined America's Role in the World}. New York: PublicAffairs, 2016.}

\item \lr{Chomsky, Noam. \textit{The New Military Humanism: Lessons from Kosovo}. Monroe, ME: Common Courage Press, 1999.}

\item \lr{Chorin, Ethan. \textit{Exit the Colonel: The Hidden History of the Libyan Revolution}. New York: PublicAffairs, 2012.}

\item \lr{Cialdini, Robert B. \textit{Influence: The Psychology of Persuasion}. Rev. ed., New York: HarperBusiness, 2006.}

\item \lr{Cohen, Florette, et al. "Fatal Attraction: The Effects of Mortality Salience on Evaluations of Charismatic, Task-Oriented, and Relationship-Oriented Leaders," \textit{Psychological Science}, Vol. 15, No. 12, 2004, pp. 846–851.}

\item \lr{Colgan, Jeff D. "Oil, Domestic Politics, and International Conflict," \textit{Energy Research \& Social Science}, Vol. 1, 2014, pp. 198–205.}

\item \lr{Comaroff, Jean and John L. Comaroff. \textit{Of Revelation and Revolution, Vol. 1: Christianity, Colonialism, and Consciousness in South Africa}. Chicago: University of Chicago Press, 1991.}

\item \lr{Commins, David. \textit{The Wahhabi Mission and Saudi Arabia}. London: I.B. Tauris, 2006.}

\item \lr{Corden, W. Max and J. Peter Neary. "Booming Sector and De-Industrialisation in a Small Open Economy," \textit{The Economic Journal}, Vol. 92, No. 368, 1982, pp. 825–848.}

\item \lr{Crawford, Neta C. and Catherine Lutz (eds.). "Costs of War Project," Watson Institute, Brown University, 2021 update.}

%% ── D ──
\item \lr{Dallaire, Roméo. \textit{Shake Hands with the Devil: The Failure of Humanity in Rwanda}. Toronto: Random House Canada, 2003.}

\item \lr{Danilevsky, Nikolay. \textit{Russia and Europe}. 1869. English trans. Stephen Woodburn. Bloomington: Slavica Publishers, 2013.}

\item \lr{Des Forges, Alison. \textit{Leave None to Tell the Story: Genocide in Rwanda}. New York: Human Rights Watch, 1999.}

\item \lr{de Waal, Alex. "US War Crimes in Somalia," \textit{New Left Review}, No. 230, 1998, pp. 131–144.}

\item \lr{DeYoung, Karen. \textit{Soldier: The Life of Colin Powell}. New York: Alfred A. Knopf, 2006.}

\item \lr{Diamond, Larry. \textit{Squandered Victory: The American Occupation and the Bungled Effort to Bring Democracy to Iraq}. New York: Times Books, 2005.}

\item \lr{di Lampedusa, Giuseppe Tomasi. \textit{The Leopard}. Trans. Archibald Colquhoun. New York: Pantheon Books, 1960.}

\item \lr{DiResta, Renée, et al. "The Tactics \& Tropes of the Internet Research Agency," \textit{New Knowledge Report for the US Senate Select Committee on Intelligence}, 2019.}

\item \lr{Dizaji, Sajjad Faraji and Peter A.G. van Bergeijk. "Potential Early Phase Success and Ultimate Failure of Economic Sanctions," \textit{Journal of Peace Research}, Vol. 50, No. 6, 2013, pp. 721–736.}

\item \lr{Duelfer, Charles. \textit{Comprehensive Report of the Special Advisor to the DCI on Iraq's WMD} (Duelfer Report). Central Intelligence Agency, 2004.}

%% ── E ──
\item \lr{Easterly, William. \textit{The White Man's Burden: Why the West's Efforts to Aid the Rest Have Done So Much Ill and So Little Good}. New York: Penguin Press, 2006.}

\item \lr{Elder, John. \textit{History of the American Presbyterian Mission in Iran, 1834–1960}. Tehran: Literature Committee of the Church Council of Iran, 1960.}

%% ── F ──
\item \lr{Fanon, Frantz. \textit{The Wretched of the Earth}. Trans. Richard Philcox. New York: Grove Press, 2004 (orig. 1961).}

\item \lr{Feinstein, Andrew. \textit{The Shadow World: Inside the Global Arms Trade}. New York: Farrar, Straus and Giroux, 2011.}

\item \lr{Ferris, Jesse. \textit{Nasser's Gamble: How Intervention in Yemen Caused the Six-Day War and the Decline of Egyptian Power}. Princeton: Princeton University Press, 2013.}

\item \lr{Frank, Andre Gunder. "The Development of Underdevelopment," \textit{Monthly Review}, Vol. 18, No. 4, 1966, pp. 17–31.}

\item \lr{Freire, Paulo. \textit{Pedagogy of the Oppressed}. Trans. Myra Bergman Ramos. New York: Continuum, 1970.}

\item \lr{Freud, Sigmund. "Beyond the Pleasure Principle" (1920), in \textit{The Standard Edition}, Vol. XVIII, London: Hogarth Press, 1955, pp. 1–64.}

\item \lr{Fukuyama, Francis. \textit{State-Building: Governance and World Order in the 21st Century}. Ithaca: Cornell University Press, 2004.}

%% ── G ──
\item \lr{Galeano, Eduardo. \textit{Open Veins of Latin America: Five Centuries of the Pillage of a Continent}. Trans. Cedric Belfrage. New York: Monthly Review Press, 1973.}

\item \lr{Garwood-Gowers, Andrew. "The Responsibility to Protect after Libya and Syria," \textit{Melbourne Journal of International Law}, Vol. 13, No. 1, 2012, pp. 594–618.}

\item \lr{Gerasimov, Valery. "The Value of Science Is in the Foresight," \textit{Military-Industrial Kurier}, 27 February 2013.}

\item \lr{Gerges, Fawaz A. \textit{ISIS: A History}. Princeton: Princeton University Press, 2016.}

\item \lr{Goldberg, Jeffrey. "The Obama Doctrine," \textit{The Atlantic}, April 2016.}

\item \lr{Gordon, Joy. \textit{Invisible War: The United States and the Iraq Sanctions}. Cambridge, MA: Harvard University Press, 2010.}

\item \lr{Gordon, Michael R. and Bernard E. Trainor. \textit{The Generals' War: The Inside Story of the Conflict in the Gulf}. Boston: Little, Brown, 1995.}

\item \lr{Gramsci, Antonio. \textit{Selections from the Prison Notebooks}. Edited and translated by Quintin Hoare and Geoffrey Nowell Smith. New York: International Publishers, 1971.}

\item \lr{Grandin, Greg. \textit{Empire's Workshop: Latin America, the United States, and the Rise of the New Imperialism}. New York: Metropolitan Books, 2006.}

\item \lr{Greenberg, Jeff, Sheldon Solomon, and Tom Pyszczynski. "Terror Management Theory of Self-Esteem and Cultural Worldviews," in \textit{Advances in Experimental Social Psychology}, Vol. 29, 1997, pp. 61–139.}

\item \lr{Guilhot, Nicolas. \textit{The Democracy Makers: Human Rights and the Politics of Global Order}. New York: Columbia University Press, 2005.}

\item \lr{Gutiérrez, Gustavo. \textit{A Theology of Liberation}. Trans. Sister Caridad Inda and John Eagleson. Maryknoll, NY: Orbis Books, 1973.}

%% ── H ──
\item \lr{Halliday, Denis. "The Impact of the UN Sanctions on the People of Iraq," \textit{Journal of Palestine Studies}, Vol. 28, No. 2, 1999, pp. 29–37.}

\item \lr{Hartung, William D. \textit{Prophets of War: Lockheed Martin and the Making of the Military-Industrial Complex}. New York: Nation Books, 2011.}

\item \lr{Hehir, Aidan. \textit{Hollow Norms and the Responsibility to Protect}. London: Palgrave Macmillan, 2019.}

\item \lr{Herman, Edward S. and Noam Chomsky. \textit{Manufacturing Consent: The Political Economy of the Mass Media}. New York: Pantheon Books, 1988; updated ed., 2002.}

\item \lr{Hochschild, Adam. \textit{King Leopold's Ghost: A Story of Greed, Terror, and Heroism in Colonial Africa}. Boston: Houghton Mifflin, 1998.}

\item \lr{Howard, Philip N. and Muzammil M. Hussain. \textit{Democracy's Fourth Wave? Digital Media and the Arab Spring}. Oxford: Oxford University Press, 2013.}

%% ── I-J ──
\item \lr{Independent International Commission on Kosovo. \textit{The Kosovo Report}. Oxford: Oxford University Press, 2000.}

\item \lr{Janis, Irving L. \textit{Victims of Groupthink}. Boston: Houghton Mifflin, 1972.}

\item \lr{Jervis, Robert. \textit{Why Intelligence Fails: Lessons from the Iranian Revolution and the Iraq War}. Ithaca: Cornell University Press, 2010.}

\item \lr{Judah, Tim. \textit{Kosovo: War and Revenge}. 2nd ed., New Haven: Yale University Press, 2002.}

%% ── K ──
\item \lr{Kahneman, Daniel. \textit{Thinking, Fast and Slow}. New York: Farrar, Straus and Giroux, 2011.}

\item \lr{Kahneman, Daniel and Jonathan Renshon. "Why Hawks Win," \textit{Foreign Policy}, No. 158, 2007, pp. 34–38.}

\item \lr{Karl, Terry Lynn. \textit{The Paradox of Plenty: Oil Booms and Petro-States}. Berkeley: University of California Press, 1997.}

\item \lr{Kenyatta, Jomo. \textit{Facing Mount Kenya}. London: Secker and Warburg, 1938.}

\item \lr{Klare, Michael T. \textit{Resource Wars: The New Landscape of Global Conflict}. New York: Metropolitan Books, 2001.}

\item \lr{Klein, Naomi. \textit{The Shock Doctrine: The Rise of Disaster Capitalism}. New York: Metropolitan Books, 2007.}

\item \lr{Kosinski, Michal, David Stillwell, and Thore Graepel. "Private Traits and Attributes Are Predictable from Digital Records," \textit{PNAS}, Vol. 110, No. 15, 2013, pp. 5802–5805.}

\item \lr{Kunda, Ziva. "The Case for Motivated Reasoning," \textit{Psychological Bulletin}, Vol. 108, No. 3, 1990, pp. 480–498.}

\item \lr{Kuperman, Alan J. "Obama's Libya Debacle," \textit{Foreign Affairs}, Vol. 94, No. 2, 2015, pp. 66–77.}

%% ── L ──
\item \lr{Lackner, Helen. \textit{Yemen in Crisis: The Road to War}. London: Verso, 2019.}

\item \lr{Lacroix, Stéphane. \textit{Awakening Islam: The Politics of Religious Dissent in Contemporary Saudi Arabia}. Cambridge, MA: Harvard University Press, 2011.}

\item \lr{Lambeth, Benjamin S. \textit{NATO's Air War for Kosovo}. Santa Monica: RAND Corporation, 2001.}

\item \lr{Landau, Jacob M. \textit{Pan-Turkism: From Irredentism to Cooperation}. 2nd ed., Bloomington: Indiana University Press, 1995.}

\item \lr{Laruelle, Marlène. \textit{Russian Nationalism: Imaginaries, Doctrines, and Political Battlefields}. London: Routledge, 2019.}

\item \lr{Lerner, Jennifer S., et al. "Effects of Fear and Anger on Perceived Risks of Terrorism," \textit{Psychological Science}, Vol. 14, No. 2, 2003, pp. 144–150.}

\item \lr{Lesch, David W. \textit{Syria: The Fall of the House of Assad}. New Haven: Yale University Press, 2012.}

\item \lr{Livingston, Steven. "Clarifying the CNN Effect," \textit{Joan Shorenstein Center Research Paper}, R-18, Harvard University, 1997.}

\item \lr{Lizza, Ryan. "The Consequentialist," \textit{The New Yorker}, 2 May 2011.}

%% ── M ──
\item \lr{Mahdavy, Hossein. "The Patterns and Problems of Economic Development in Rentier States: The Case of Iran," in M.A. Cook (ed.), \textit{Studies in the Economic History of the Middle East}. London: Oxford University Press, 1970, pp. 428–467.}

\item \lr{Makiya, Kanan. \textit{Republic of Fear: The Politics of Modern Iraq}. Updated ed., Berkeley: University of California Press, 1998.}

\item \lr{Makiya, Kanan. \textit{Cruelty and Silence: War, Tyranny, Uprising, and the Arab World}. New York: W.W. Norton, 1993.}

\item \lr{Martín-Baró, Ignacio. \textit{Writings for a Liberation Psychology}. Cambridge, MA: Harvard University Press, 1994.}

\item \lr{Mearsheimer, John J. and Stephen M. Walt. \textit{The Israel Lobby and U.S. Foreign Policy}. New York: Farrar, Straus and Giroux, 2007.}

\item \lr{Mignolo, Walter D. \textit{The Darker Side of the Renaissance}. 2nd ed., Ann Arbor: University of Michigan Press, 2003.}

\item \lr{Milgram, Stanley. \textit{Obedience to Authority: An Experimental View}. New York: Harper \& Row, 1974.}

\item \lr{Mill, John Stuart. "A Few Words on Non-Intervention," \textit{Fraser's Magazine}, December 1859.}

\item \lr{Mills, C. Wright. \textit{The Sociological Imagination}. Oxford: Oxford University Press, 1959.}

\item \lr{Mitchell, Lincoln A. \textit{The Color Revolutions}. Philadelphia: University of Pennsylvania Press, 2012.}

\item \lr{Mueller, John and Karl Mueller. "Sanctions of Mass Destruction," \textit{Foreign Affairs}, Vol. 78, No. 3, 1999, pp. 43–53.}

\item \lr{Muttitt, Greg. \textit{Fuel on the Fire: Oil and Politics in Occupied Iraq}. London: Bodley Head, 2011.}

%% ── N ──
\item \lr{Nasr, Vali. \textit{The Shia Revival: How Conflicts Within Islam Will Shape the Future}. New York: W.W. Norton, 2006.}

\item \lr{Nickerson, Raymond S. "Confirmation Bias: A Ubiquitous Phenomenon in Many Guises," \textit{Review of General Psychology}, Vol. 2, No. 2, 1998, pp. 175–220.}

\item \lr{Norton, Augustus Richard. \textit{Hezbollah: A Short History}. Princeton: Princeton University Press, 2007.}

\item \lr{Nye, Joseph S. Jr. \textit{Soft Power: The Means to Success in World Politics}. New York: PublicAffairs, 2004.}

%% ── O-P ──
\item \lr{Ouimet, Matthew J. \textit{The Rise and Fall of the Brezhnev Doctrine in Soviet Foreign Policy}. Chapel Hill: University of North Carolina Press, 2003.}

\item \lr{Pariser, Eli. \textit{The Filter Bubble: What the Internet Is Hiding from You}. New York: Penguin Press, 2011.}

\item \lr{Phillips, Christopher. \textit{The Battle for Syria: International Rivalry in the New Middle East}. New Haven: Yale University Press, 2016.}

\item \lr{Power, Samantha. \textit{A Problem from Hell: America and the Age of Genocide}. New York: Basic Books, 2002.}

\item \lr{Prashad, Vijay. \textit{The Darker Nations: A People's History of the Third World}. New York: The New Press, 2007.}

\item \lr{Prunier, Gérard. \textit{The Rwanda Crisis: History of a Genocide}. New York: Columbia University Press, 1997.}

\item \lr{Putnam, Robert D. \textit{Making Democracy Work: Civic Traditions in Modern Italy}. Princeton: Princeton University Press, 1993.}

%% ── R ──
\item \lr{Rashid, Ahmed. \textit{Pakistan on the Brink}. New York: Viking, 2012.}

\item \lr{Ricks, Thomas E. \textit{Fiasco: The American Military Adventure in Iraq}. New York: Penguin Press, 2006.}

\item \lr{Rid, Thomas. \textit{Active Measures: The Secret History of Disinformation and Political Warfare}. New York: Farrar, Straus and Giroux, 2020.}

\item \lr{Rid, Thomas. \textit{Cyber War Will Not Take Place}. London: Hurst \& Company, 2013.}

\item \lr{Robinson, Piers. \textit{The CNN Effect: The Myth of News, Foreign Policy and Intervention}. London: Routledge, 2002.}

\item \lr{Rodrik, Dani. "Goodbye Washington Consensus, Hello Washington Confusion?" \textit{Journal of Economic Literature}, Vol. 44, No. 4, 2006, pp. 973–987.}

\item \lr{Roosevelt, Eleanor. \textit{This Is My Story}. New York: Harper \& Brothers, 1937.}

\item \lr{Ross, Michael L. \textit{The Oil Curse: How Petroleum Wealth Shapes the Development of Nations}. Princeton: Princeton University Press, 2012.}

\item \lr{Rozin, Paul and Edward B. Royzman. "Negativity Bias, Negativity Dominance, and Contagion," \textit{Personality and Social Psychology Review}, Vol. 5, No. 4, 2001, pp. 296–320.}

%% ── S ──
\item \lr{Said, Edward W. \textit{Orientalism}. New York: Pantheon Books, 1978.}

\item \lr{Sakwa, Richard. \textit{Frontline Ukraine: Crisis in the Borderlands}. London: I.B. Tauris, 2015.}

\item \lr{Salisbury, Peter. "Yemen: National Chaos, Local Order," \textit{Chatham House Report}, December 2017.}

\item \lr{Sanger, David E. \textit{Confront and Conceal: Obama's Secret Wars and Surprising Use of American Power}. New York: Crown, 2012.}

\item \lr{Sarotte, Mary Elise. \textit{Not One Inch: America, Russia, and the Making of Post-Cold War Stalemate}. New Haven: Yale University Press, 2021.}

\item \lr{Scahill, Jeremy. \textit{The Assassination Complex: Inside the Government's Secret Drone Warfare Program}. New York: Simon \& Schuster, 2016.}

\item \lr{Scott, James C. \textit{Weapons of the Weak: Everyday Forms of Peasant Resistance}. New Haven: Yale University Press, 1985.}

\item \lr{Seligman, Martin E.P. \textit{Helplessness: On Depression, Development, and Death}. San Francisco: W.H. Freeman, 1975.}

\item \lr{Seligman, Martin E.P. and Steven F. Maier. "Failure to Escape Traumatic Shock," \textit{Journal of Experimental Psychology}, Vol. 74, No. 1, 1967, pp. 1–9.}

\item \lr{Shaffer, Brenda. \textit{Borders and Brethren: Iran and the Challenge of Azerbaijani Identity}. Cambridge, MA: MIT Press, 2002.}

\item \lr{Sharot, Tali. \textit{The Optimism Bias}. New York: Pantheon Books, 2011.}

\item \lr{Shuster, Simon. \textit{The Showman: Inside the Invasion That Shook the World and Made a Leader of Volodymyr Zelensky}. New York: William Morrow, 2024.}

\item \lr{Singer, P.W. \textit{Corporate Warriors: The Rise of the Privatized Military Industry}. Ithaca: Cornell University Press, 2003.}

\item \lr{Stearns, Jason K. \textit{Dancing in the Glory of Monsters: The Collapse of the Congo and the Great War of Africa}. New York: PublicAffairs, 2011.}

\item \lr{Stiglitz, Joseph E. \textit{Globalization and Its Discontents}. New York: W.W. Norton, 2002.}

\item \lr{Stiglitz, Joseph E. and Linda J. Bilmes. \textit{The Three Trillion Dollar War: The True Cost of the Iraq Conflict}. New York: W.W. Norton, 2008.}

\item \lr{Sunstein, Cass R. "The Law of Group Polarization," \textit{Journal of Political Philosophy}, Vol. 10, No. 2, 2002, pp. 175–195.}

%% ── T ──
\item \lr{Tajfel, Henri and John C. Turner. "An Integrative Theory of Intergroup Conflict," in Austin and Worchel (eds.), \textit{The Social Psychology of Intergroup Relations}. Monterey, CA: Brooks/Cole, 1979, pp. 33–47.}

\item \lr{Taleb, Nassim Nicholas. \textit{The Black Swan: The Impact of the Highly Improbable}. 2nd ed., New York: Random House, 2010.}

\item \lr{Taşpınar, Ömer. "Turkey's Middle East Policies: Between Neo-Ottomanism and Kemalism," \textit{Carnegie Endowment Papers}, No. 10, 2008.}

\item \lr{Tibi, Bassam. \textit{Arab Nationalism: Between Islam and the Nation-State}. 3rd ed., London: Macmillan, 1997.}

\item \lr{Todorova, Maria. \textit{Imagining the Balkans}. Oxford: Oxford University Press, 1997.}

\item \lr{Turner, John C., et al. \textit{Rediscovering the Social Group: A Self-Categorization Theory}. Oxford: Basil Blackwell, 1987.}

%% ── V-W ──
\item \lr{Volcker, Paul A., Richard J. Goldstone, and Mark Pieth. \textit{The Management of the United Nations Oil-for-Food Programme}. Independent Inquiry Committee, 2005.}

\item \lr{Volkan, Vamık D. \textit{Bloodlines: From Ethnic Pride to Ethnic Terrorism}. New York: Farrar, Straus and Giroux, 1997.}

\item \lr{Volkan, Vamık D. "Transgenerational Transmissions and Chosen Traumas," \textit{Group Analysis}, Vol. 34, No. 1, 2001, pp. 79–97.}

\item \lr{Volkan, Vamık D. \textit{Killing in the Name of Identity}. Charlottesville: Pitchstone Publishing, 2006.}

\item \lr{Wallerstein, Immanuel. \textit{The Modern World-System I}. New York: Academic Press, 1974.}

\item \lr{Wallerstein, Immanuel. \textit{The Decline of American Power}. New York: The New Press, 2003.}

\item \lr{Walsh, Lawrence E. \textit{Iran-Contra: The Final Report}. New York: Times Books, 1994.}

\item \lr{Walzer, Michael. \textit{Just and Unjust Wars}. 5th ed., New York: Basic Books, 2015.}

\item \lr{wa Thiong'o, Ngũgĩ. \textit{Decolonising the Mind: The Politics of Language in African Literature}. London: James Currey, 1986.}

\item \lr{Wehrey, Frederic. \textit{The Burning Shores: Inside the Battle for the New Libya}. New York: Farrar, Straus and Giroux, 2018.}

\item \lr{Westad, Odd Arne. \textit{The Global Cold War: Third World Interventions and the Making of Our Times}. Cambridge: Cambridge University Press, 2005.}

\item \lr{Wheeler, Nicholas J. \textit{Saving Strangers: Humanitarian Intervention in International Society}. Oxford: Oxford University Press, 2000.}

\item \lr{Whitlock, Craig. \textit{The Afghanistan Papers: A Secret History of the War}. New York: Simon \& Schuster, 2021.}

\item \lr{Williams, Paul D. \textit{War and Conflict in Africa}. 2nd ed., Cambridge: Polity Press, 2016.}

%% ── Z ──
\item \lr{Zetter, Kim. \textit{Countdown to Zero Day: Stuxnet and the Launch of the World's First Digital Weapon}. New York: Crown, 2014.}

\item \lr{Zifcak, Spencer. "The Responsibility to Protect after Libya and Syria," \textit{Melbourne Journal of International Law}, Vol. 13, No. 1, 2012, pp. 59–93.}

\item \lr{Zimbardo, Philip G. \textit{The Lucifer Effect: Understanding How Good People Turn Evil}. New York: Random House, 2007.}

\item \lr{Zirinsky, Michael P. "Render Therefore unto Caesar the Things Which Are Caesar's: American Presbyterian Educators and Reza Shah," \textit{Iranian Studies}, Vol. 26, No. 3-4, 1993, pp. 337–356.}

\end{enumerate}


%% ================================================================
\section*{پ) اسناد رسمی و بین‌المللی}
\addcontentsline{toc}{section}{اسناد رسمی و بین‌المللی}
%% ================================================================

\begin{enumerate}[label={[\arabic*]}, leftmargin=2cm, itemsep=6pt, resume]

\item \lr{Bush, George H.W. "Address Before a Joint Session of Congress on the State of the Union," January 29, 1991. \textit{Public Papers of the Presidents}.}

\item \lr{Central Intelligence Agency. Declassified Documents: Iran 1953 Coup. Released 2017.}

\item \lr{Constitutive Act of the African Union, Article 4(h), adopted in Lomé, Togo, 11 July 2000.}

\item \lr{Eisenhower, Dwight D. "Farewell Address to the Nation," January 17, 1961.}

\item \lr{International Rescue Committee. "Mortality in the Democratic Republic of Congo: An Ongoing Crisis," 2007.}

\item \lr{Iraq Body Count. \url{https://www.iraqbodycount.org/}. Accessed 2024.}

\item \lr{Khomeini, Ruhollah. "New Year's Message," 21 March 1980. In \textit{Sahifeh-ye Imam}, Vol. 12.}

\item \lr{Mwatana for Human Rights. \textit{Day of Judgment: The Role of the US and Europe in Civilian Death, Destruction, and Trauma in Yemen}. 2019.}

\item \lr{NATO. "Operation Unified Protector: Final Mission Stats," 2 November 2011.}

\item \lr{Oxfam International. "Inequality Kills," Davos Report, January 2022.}

\item \lr{Putin, Vladimir. "Address by the President of the Russian Federation," 24 February 2022.}

\item \lr{Special Inspector General for Afghanistan Reconstruction (SIGAR). \textit{What We Need to Learn}. Washington, DC, 2021.}

\item \lr{Special Inspector General for Iraq Reconstruction (SIGIR). \textit{Learning from Iraq: A Final Report}. Washington, DC, 2013.}

\item \lr{Syrian Network for Human Rights. "Russia Has Killed More Syrian Civilians Than ISIS," Report, 2016.}

\item \lr{Syrian Observatory for Human Rights. Cumulative Death Toll Report, 2023.}

\item \lr{UNESCO. \textit{Education for All Global Monitoring Report: The Hidden Crisis}. Paris: UNESCO, 2011.}

\item \lr{UNICEF. "Child and Maternal Mortality Survey," 1999.}

\item \lr{United Nations Security Council. Resolution 660 (1990), S/RES/660.}

\item \lr{United Nations Security Council. Resolution 678 (1990), S/RES/678.}

\item \lr{United Nations Security Council. Resolution 794 (1992), S/RES/794.}

\item \lr{United Nations Security Council. Resolution 1368 (2001), S/RES/1368.}

\item \lr{United Nations Security Council. Resolution 1973 (2011), S/RES/1973.}

\item \lr{US Congress. Joint Inquiry into Intelligence Community Activities, "28 Pages," declassified July 2016.}

\item \lr{Watson Institute, Brown University. "Costs of War Project," 2021. \url{https://watson.brown.edu/costsofwar/}.}

\item \lr{World Bank. \textit{The Toll of War: The Economic and Social Consequences of the Conflict in Syria}. 2017.}

\item \lr{World Bank. "Afghanistan Development Update," October 2022.}

\end{enumerate}

\cleardoublepage